### 2020-08-20 auxilary_scripts/Finrep_data_comparison.Rmd
add import and process functions for oracle finrep data

```{r load_libraries}

library(devtools)

load_all()

library(tidyverse)

```

```{r Import_data}

short_df = import.boi.finrep.data()

long_df = import.boi.oracle.finrep.data()

long_df = long_df %>%
  filter(report_period == "quarterly")

```

```{r gazit_glob_comparison}

X = full_join(
  long_df %>% 
    filter(tase_id == 126) %>% 
    filter(year(date_yearmon) %in% c(2000,2007,2010)) %>% 
    mutate(date_yearqtr = as.yearqtr(date_yearmon)) %>% 
    mutate(total_assets = as.numeric(total_assets) / 1000) %>% 
    select(tase_id, date_yearmon,date_yearqtr, total_assets),
  short_df %>% 
  filter(TASE_ID == 126) %>% 
  filter(year(Date) %in% c(2000,2007,2010)) %>% 
  mutate(Total_Assets = as.numeric(Total_Assets)) %>% 
  select(TASE_ID,Date, Total_Assets),
  by = c("total_assets" = "Total_Assets")
  ) %>% 
  select(date_yearmon,date_yearqtr,Date,total_assets)

X %>% 
  mutate(Month = month(date_yearmon)) %>% 
  arrange(Date) %>% 
  select(Month, everything())

```