### 2020-08-21 auxilary_scripts/Finrep_data_comparison.Rmd
add oracle and nimrod comparison

----
title: Financial reports data  - comparison of sources
----

The main conclusion is that there are differences between Oracle, Mos and Nimrod's data. However the differences are not systematic, for the same tase_id, variable and
date there are sometimes identical observations and sometimes a diffrerent ones.

```{r load_libraries}

library(devtools)

load_all()

library(tidyverse)

```

```{r Import_data}

secs_catalog = read_csv(paste0(
  "C:\\Users\\internet\\OneDrive - Bank Of Israel",
  "\\Data\\TASE liquidity\\Secs_Catalog.csv")) %>% 
  mutate(across(c("tase_id","sec_id"), as.character))

mos_df = import.boi.finrep.data() %>% 
  rename_all(tolower) %>% 
  rename(date_yearqtr = date) %>% 
  rename(total_net_profit = net_profit)

oracle_df  = import.boi.oracle.finrep.data() %>%
  filter(reporting_period == "quarterly")

nimrod_df = import.nimrod.stata.df(paste0(
  file.path(Sys.getenv("USERPROFILE"),fsep = "\\"),
  "\\OneDrive - Bank Of Israel\\Data\\",
  "TASE liquidity\\Stata files",
  "\\TASE panel.dta")
  ) %>% 
  rename(date_yearqtr = date)

```

\section{Oracle vs Mos comparison}

```{r oracle_mos_df}

oracle_mos_diff_df = full_join(
  oracle_df %>% 
    select(intersect(names(.), names(mos_df))) %>% 
    mutate(across(-c("tase_id", "date_yearqtr"), ~ as.numeric(.) * 10 ^ (-3))) %>% 
    pivot_longer(-c("tase_id", "date_yearqtr"),
                 names_to = "feature", values_to = "oracle"),
  mos_df %>% 
    select(intersect(names(.), names(oracle_df))) %>% 
    mutate(across(-c("tase_id", "date_yearqtr"),as.numeric)) %>%  
    pivot_longer(-c("tase_id", "date_yearqtr"),
                 names_to = "feature", values_to = "mos"),
  by = c("tase_id","date_yearqtr","feature"))%>% 
  rowwise() %>% 
  mutate(diff = abs((oracle - mos) / mean(c(oracle,mos))) * 100) %>% 
  ungroup() %>% 
  arrange(desc(diff))

```

There are some differences between oracle and mos databases.
In the case of Aura company oracle database is correct
(for example the comparison of 2019Q1 profit). For Brand company mos is correct.

\section{Oracle vs Nimrod comparison}

```{r oracle_nimrod_df}

oracle_nimrod_diff_df = full_join(
  oracle_df %>% 
    select(intersect(names(.), names(nimrod_df))) %>% 
    mutate(across(-c("tase_id", "date_yearqtr"), ~ as.numeric(.) * 10 ^ (-3))) %>% 
    pivot_longer(-c("tase_id", "date_yearqtr"),
                 names_to = "feature", values_to = "oracle"),
  nimrod_df %>% 
    select(intersect(names(.), names(oracle_df))) %>% 
    mutate(across(-c("tase_id", "date_yearqtr"), ~ as.numeric(.) * 10 ^ (-3))) %>% 
    mutate(across(-c("tase_id", "date_yearqtr"),as.numeric)) %>%  
    pivot_longer(-c("tase_id", "date_yearqtr"),
                 names_to = "feature", values_to = "nimrod"),
  by = c("tase_id","date_yearqtr","feature"))%>% 
  rowwise() %>% 
  mutate(diff = abs((oracle - nimrod) / mean(c(oracle,nimrod))) * 100) %>% 
  ungroup() %>% 
  arrange(desc(diff))

discrep_df = oracle_nimrod_diff_df %>% 
  filter(diff > 5) %>% 
  left_join(secs_catalog[,c("tase_id","comp_name_heb")])

```

There are